### 2020-10-31 Modules/HPBW_apndx.tex
Initial Overleaf Import

We now compute the squared norm of \eqref{eq_Hdphi0}, equate it to $1/2$ and compute the value of $N\dTheta/2$ for large $N$. To this end, we denote $\gamma\triangleq\dTheta/2$, giving rise to
\begin{equation*}
    \resizebox{1\linewidth}{!}{
        \begin{split}
            \abs{\Hr_{\dTheta,\dPhi=0,r}\rBrace{\omega}}^{2}&=
             \frac{\rBrace{1-r}^{2}\Dp{\gamma,N}{2}}{\abs{\exp\rBrace{j(N-1)\gamma}-r\D{\gamma}{N}}^{2}}
             \\&
             \overset{N\gg1}{\approx}\frac{\rBrace{1-r}^{2}\Dp{\gamma,N}{2}}{1-2r\cos{\rBrace{N\gamma}}\D{\gamma}{N}+r^{2}\Dp{\gamma,N}{2}}.
        \end{split}
    }
\end{equation*}
Equating to $1/2$, leads to
\begin{equation*}
    % \resizebox{1\linewidth}{!}{
        \begin{split}
            \rBrace{r^{2}-4r+2}\Dp{\gamma,N}{2}+2r\cos{\rBrace{N\gamma}}\D{\gamma}{N}-1 = 0.
        \end{split}
    % }
\end{equation*}
Since for large $N$, the mainlobe beamwidth goes to zero, we approximate $\sin(\gamma)$ with $\gamma$. Also defining $x\triangleq{}N\gamma=N\dTheta/2$ we obtain \eqref{eq_HPBW}.