### 2020-10-31 Modules/IEEE_journal_paper_Intro.tex

general field of array processing has been thoroughly studied throughout several decades.
The array sensors' spatial diversity enables the extraction of spatial information about impinging signals, thus laying the ground for wide range of applications, such as localizing a transmitting source \cite{skolnik2008radar,long2019}, blindly separating mixtures of impinging signals \cite{comon1994independent}, improving signal to noise ratio (SNR) \cite{verdu1998multiuser}, and many more. 
\par A uniform linear array (ULA) has always been a point of interest, due to its simplicity of analysis \cite{van2004optimum,benesty2018}. 
The array size and the number of its elements ($N$) has significant influence on the obtained array performance, such as SNR improvement, spatial separation capabilities and its spatial response's degrees of freedom (DOF). For example, the influential MUSIC algorithm \cite{schmidt1986multiple} enables the localization of signals arriving from up to $N-1$ distinctive directions of arrival (DOA), by projecting the array manifold onto the noise subspace.
\par In pursuit of spatial performance improvement, namely higher spatial separation and selectivity of arriving signals, many approaches were suggested.  
One approach, commonly referenced as ``virtual arrays'' \cite{pal2010nested,chevalier2005virtual,dogan1995applications} deals with the extraction of samples originated in sensors that do not really exist by using high (higher than 2) order statistics and manipulating multiple statistical cross-terms in order to estimate statistical characteristics of signals impinging in missing sensors.
Using a similar approach, the $2q$-MUSIC algorithm \cite{chevalier2006high}, enables the use of $N^{2q}$ ``virtual elements'', by calculating the $q$'th order statistics.
Another approach, involving different array geometries, examined minimum redundancy arrays \cite{moffet1968minimum,pillai1985new,pillai1987statistical,Kupershtein2013}, aiming to reduce the spatial ambiguity. The basic concept was minimization of the inter-element spacing redundancy in order to increase the overall resolution. Adaptive processing schemes \cite{frost1972algorithm,manolakis2000statistical}, being a wide and active  research area, were also suggested trying to adaptively estimate and suppress the noise component in impinging signals by minimization of the receiver's output energy with some constraints.
\par Pursuing other approaches to improve the array's spatial performance, ULA spatial array processing analogy to finite impulse response (FIR) \cite{van1988beamforming} and the infinite impulse response (IIR) superior performance (e.g. narrower transition regions and higher sidelobes attenuation) gave rise to the question ``what are the equivalent spatial domain processing methods which will be analogous to temporal IIR filtering?''
\par Achieving spatial IIR response has also motivated other works.
In \cite{wen2013extending} two methods were considered.
The first one was to estimate the time of arrival (TOA) difference between two consecutive sensors and to synthetically generate the recursive part of the IIR filter, entirely in the time-domain. The second approach suggested to consider overlapping subsets of one large ULA as finite approximation to an infinite array. 
Surely, the former approach heavily relies on the accuracy of the delay estimation and the latter approach does not achieve a recursive spatial response. In both cases, there is no true spatial feedback between the array and the source of interest.
Also, ultra-wideband (UWB) filters, which sample spatial snapshots of the signal and recursively process it in temporal domain were designed in \cite{bruton1983highly}, using the $2D$ spatio-temporal plane wave representation as a straight line angled according to the DOA.
\par In this contribution, we present a low-complexity sensor array processing approach which achieves the desired spatial domain exclusive IIR-like beampattern, while avoiding any temporal processing of the signal.
To this end, we arbitrarily choose to formulate the problem in the context of a localization problem, hence our goal is to estimate the direction and the range of some target of interest. 
\par The novelty, compared to traditional array processing, is the incorporation of spatial feedback, which we prove to be the spatial domain equivalent of temporal domain IIR filtering.
Assuming the target of interest has a mirror-like behaviour (i.e., reflects its impinging signals), the spatial feedback between the array and the target is created by continuously re-transmitting a synthesized version of the impinging signal (and its reflections) to the target.
Note that the initial stimulus can be generated by the target or the array itself. In the text to follow, we assume this is the latter. 
Furthermore, as opposed to the passive target case (i.e., a target which merely reflects the impinging signal), one may consider a cooperative target, which receives, enhances and re-transmits the signal back to the array. 
\par The outline of this paper is as follows. We first formulate the classic spatial beamforming setup in Sec.~\ref{sec:setup}. Then, in Sec.~\ref{sec_introduceFeedback}, we propose our novel feedback-based architecture, and formulate its spatial response.
Searching for localization performance maximization, Sec.~ \ref{sec_FIM} discusses the information-theory related considerations for the array configuration, utilizing the Fisher Information Matrix (FIM) in the context of the target's range and DOA estimation.
In Sec.~\ref{sec_Performance}, we evaluate some key features of the proposed beamforming with feedback. Specifically, we compute the array beamwidth, its peak-to-sidelobe ratio and the array directivity, showing significant improvement compared to traditional beamforming without spatial feedback. 
In Sec.~\ref{sec_app}, we simulate the proposed processing scheme, and emphasize its sensitivity to range errors. We then suggest a strategy which mitigates this sensitivity. Finally, concluding remarks are given in Sec.~\ref{sec_conclusions}.
\section{Beamforming}\label{sec:setup}
In this section, a brief overview on ULA beamforming and its resemblance to FIR filter design are presented.
Consider an $N$-element ULA with inter-element spacing $\delta$, where its $n$'th sensor is positioned at $\vpi{n},\;\text{for}\; n=0,\ldots,N-1$. We set $\vpi{0}$ as the axis reference point and assume that the target of interest is positioned at $\vpt$.
Focusing on a far field localization problem, DOA and target range are to be estimated. Without loss of generality, aiming to simplify the exposition, we assume $2\text{D}$ planar problem, where DOA is described by a single angle $\thetaD$, measured from the array's broadside.
For simplicity, we assume an anechoic environment, an array of identical omni-directional sensors and a stationary target of interest.
Inspired by radar based applications, we also place a transmitter at $\vpi{0}$, assuming the transmitted signal $s$, reflects back from the target and re-impinges the array, with total time delay of $\tau_{\text{pd}}=2R/c$ seconds, where $R = \norm{\vpt-\vpi{0}}$ is the target's range and $c$ represents the propagation velocity of the signal in the medium.
In this this work, we use the following conventions; vectors and matrices are denoted by lower and upper case bold italic letters respectively. 
$v_{i}$ and $A_{ij}$ are the $i$-th element of the vector $\vecnot{v}$ and the $i,j$-th element of the matrix $\vecnot{A}$.
Also, $\vecnot{A}^{T}$, $\vecnot{A}^{*}$ and $\vecnot{A}^{H}$ are the transpose, conjugate and hermitian transpose of the matrix $\vecnot{A}$ respectively.
\par Let $x_{n}(t)$ be the measured signal at the $n$'th sensor
\begin{equation}
x_{n}(t) = gs\Brack{t-\tau_{pd}-\tau_{n}},
\label{eqn:noFeedbackULA_singleSensor_temporal}
\end{equation}
where $\tau_{n}=n\delta\cos\Brack{\thetaD}/c$ represents the time difference of arrival between the $n$'th sensor and the reference sensor and $g$, being a scalar in anechoic environment, is the channel's gain, related to both propagation and the target's radar cross section (RCS).
Defining $\vecnot{x}\rBrace{t}\triangleq\vBrace{x_{0}\rBrace{t}\hdots{}x_{N-1}\rBrace{t}}^{T}$ and its Fourier transform, $\vecnot{\F{x}}\rBrace{\omega}\triangleq\vBrace{X_{0}\rBrace{\omega},\hdots,X_{N-1}\rBrace{\omega}}^{T}$, one may write 
\[
\vecnot{\F{x}}\rBrace{\omega}=g\vecnot{d}\rBrace{\omega}\F{s}\rBrace{\omega}\exp{\rBrace{-j\omega\tau_{pd}}}
\]
where $S\rBrace{\omega}$ is the Fourier transform of $s\rBrace{t}$ and $\vdO$ denotes the steering vector whose $n$'th element is
\begin{equation}
    \label{eq:d}
    d_{n}\rBrace{\omega} = \exp{\rBrace{-j\omega\tau_{n}}}.
\end{equation}
Denoting the beamformer's weights as $\vBeta\rBrace{\omega}$ and the beamformer's output as $z$, we express the latter in frequency domain
\begin{equation}
    \label{eq_Z}
    Z\rBrace{\omega} = g\vecnot{\beta}^{H}\rBrace{\omega}\vecnot{d}\rBrace{\omega}\F{s}\rBrace{\omega}\exp{\rBrace{-j\omega\tau_{pd}}}.
\end{equation}
Defining the electric phase to be
\begin{equation}\label{eq:thetaULA}
\theta=\omega{\delta\cos\Brack{\thetaD}}/{c},
\end{equation}
we rewrite \eqref{eq_Z} as 
\[
Z\rBrace{\omega} = g\F{s}\rBrace{\omega}\exp{\rBrace{-j\omega\tau_{pd}}}\sum_{n=0}^{N-1}\beta^{*}_{n}\rBrace{\omega}\exp\Brack{-jn\theta},
\]
hence in the ULA case, aiming for a desired spatial response, the weights vector $\vecnot{\beta}\rBrace{\omega}$ configuration is mathematically equivalent to FIR filter design~\cite{van1988beamforming,benesty2018}. 
\ifdefined\useOmega \else 
Assuming narrowband stimuli signals, we suppress $\omega$ dependency in the notation throughout the rest this paper, where possible.
\fi
\par In standard radar signal processing schemes, a waveform is transmitted to, and reflected from the target of interest. Then, the reflected signal is processed by the radar reception array in order to estimate the target's dynamics (e.g., DOA, range, velocity etc.). 
As opposed to the standard scheme, we suggest a continuous re-transmission of the signal and its echoes back to the platform, generating a spatial feedback loop between the array and the target.
Another deviation from traditional radar processing, used to simplify the exposition, is using continuous-wave (CW) stimuli, rater than using pulse based signals. 
% The suggested structure is elaborated in Sec.~\ref{sec_introduceFeedback} and shown to be equivalent to IIR filter design.
\section{Feedback Beamforming}
\label{sec_introduceFeedback}
In this section, a feedback-based architecture is proposed for spatial signal processing.
Inspired by time domain ``Direct form II'' IIR filter design (see Fig.~\ref{fig_IIRBasicArch}), 
we propose to use the same concept in the spatial domain.
\begin{figure}[t!]
    \begin{center}
        \begin{overpic}[width=0.6\linewidth, 
        % grid, 
        tics=10,trim=0 0 0 0]{./Media/BASIC_IIR_FILTER_ARCH.png}
            \put (60, 50){\footnotesize{$\beta_{0}$}}
            \put (60, 30){\footnotesize{$\beta_{1}$}}
            \put (60, 10){\footnotesize{$\beta_{2}$}}
            \put (36, 30){\footnotesize{$\alpha_{1}$}}
            \put (36, 10){\footnotesize{$\alpha_{2}$}}
            \put (10, 50){\footnotesize{$x$}}
            \put (85, 50){\footnotesize{$y$}}
            \put (47.5, 34.5){\footnotesize{$z^{-1}$}}
            \put (47.5, 14.5){\footnotesize{$z^{-1}$}}
        \end{overpic}
    \end{center}
    \caption{Direct form II $2^{nd}$ order IIR architecture.}
    \label{fig_IIRBasicArch}
\end{figure}
The suggested feedback beamformer (FB) architecture, where the output signal ($z$) is synthesized using weights $\vBeta$ and the weights $\vAlpha$ synthesize the feedback transmission ($\Tx$), is presented in Fig.~\ref{fig:Proposed_spatialIIR_ARCH}. The beamformer's output and the feedback signal are synthesized with two independently configured beamformers, $s$ is the system's stimulus and an additive noise (n) is assumed at the array's output.
Also, the FB block is marked (dashed line) for later use.
Note that setting $\vAlpha=\vecnot{0}$ (i.e., cancelling the feedback) degenerates the system to a plain delay-and-sum (DS) beamformer.
\begin{figure}[t!]
    \begin{center}
        \begin{overpic}[width=0.95\linewidth, 
        % grid, 
        tics=10,trim={0 0 0 0}]{./Media/SpatialIIR-diagram/SpatialIIR_VER8.png}
            \put (14, 39){\footnotesize{$\sum_{n=0}^{N-1}\beta^{*}_{n}x_{n}$}}
            \put (54, 39){\footnotesize{$\sum_{n=0}^{N-1}\alpha^{*}_{n}x_{n}$}}
            \put (24.25, 50){\footnotesize{$x_{0}$}}
            \put (35, 50){\footnotesize{$x_{1}$}}
            \put (55, 50){\footnotesize{$x_{N-1}$}}
            \put (30.5, 66){\footnotesize{$\delta$}}
            \put (89, 96){\footnotesize{$\vpt$}}
            \put (58, 58){\footnotesize{$\vpi{N-1}$}}
            \put (37, 58){\footnotesize{$\vpi{1}$}}
            \put (26.5, 58){\footnotesize{$\vpi{0}$}}
            \put (41.5, 64.5){\footnotesize{$\theta_{g}$}}
            \put (19, 11.25){\large{$+$}}
            \put (61.25, 27.25){\large{$+$}}
            \put (47, 11.75){$s$}
            \put (35.25, 11.75){$n$}
            \put (21,4){$z$}
            \put (63.5,14){$\Tx$}
            \put (1, 51){$\text{FB}_{\vAlpha,\vBeta}$}
        \end{overpic}
    \end{center}
    \caption{
    The proposed feedback beamformer.
    The spatial feedback is obtained by continuous re-transmission of $\Tx$ to the target at $\vpt$.
    We designate the feedback beamformer block (dashed line) for later use.
    }
    \label{fig:Proposed_spatialIIR_ARCH}
\end{figure}
\subsection*{Obtained spatial response}
\input{./Modules/Spatial_IIR-MathDerivations.tex}
\section{Fisher Information Matrix}
\label{sec_FIM}
\input{./Modules/estimation_CRLB_appendix.tex}
\section{Performance Analysis}
\label{sec_Performance}
In this section we analyze the suggested FB (see Fig.~\ref{fig:Proposed_spatialIIR_ARCH}), considering some fundamental properties which are commonly used to asses array performance: Beamwidth, peak-to-sidelobe level, and directivity. Each property is then compared to traditional passive ULAs, showing that significantly improved performances are obtainable with spatial feedback integration.  
\subsection*{Error terms}
In the absence of accurately known parameters, we denote $\hat{\phi},\hat{\theta}$ to be the range and DOA related phase estimates respectively.
Then, using the same weights as in \eqref{eq:alpha_beta_opt} for both the feedback and output synthesis gives rise to
\begin{equation}\label{eq:alpha_beta_hat}
\vBetaCFB=\vAlphaCFB=\frac{\vdHatC\exp\rBrace{j\hat{\phi}}}{\hat{g}\norm{\vdHat}^2},
\end{equation}
introducing the estimated steering vector 
\begin{equation}\label{eq:d_hat}
\vdHat=\vBrace{1,\exp(-\hat\theta),\ldots,\exp(-(N-1)\hat\theta)}^T.
\end{equation}
Plugging \eqref{eq:alpha_beta_hat} into \eqref{eqn:GeneralFeedbackTransferFunction}, results in
\begin{equation}\label{eq:SF_CB}
    \resizebox{.894\linewidth}{!}{
        \begin{split}
            \HbaFB=\frac{r\D{\dTheta/2}{N}\exp\rBrace{-j\rBrace{\dPhi+(N-1)\dTheta/2}}}{1-r\D{\dTheta/2}{N}\exp\rBrace{-j\rBrace{\dPhi+(N-1)\dTheta/2}}}
        \end{split}
    }
\end{equation}
where \[
\D{x}{N}\triangleq\frac{1}{N}\frac{\sin\rBrace{Nx}}{\sin\rBrace{x}}
\]
is the normalized Dirichlet kernel and
we define the DOA, range and gain error terms 
\[
\dTheta\triangleq\theta\omegaB-\hat{\theta}\omegaB,\ \dPhi\triangleq\phi\omegaB-\hat{\phi}\omegaB,\ 
r\triangleq g/\hat{g},
\]
respectively.
\ifdefined\useOmega\else
We then define four fundamental scenarios:
\fi
\begin{itemize}
    \item{\makebox[.35\linewidth]{Perfect alignment \hfill} $\rBrace{\dTheta=0\ , \dPhi=0},$}
    \item{\makebox[.35\linewidth]{Steer error \hfill} $\rBrace{\abs{\dTheta}>0\ , \dPhi=0},$}
    \item{\makebox[.35\linewidth]{Range error \hfill} $\rBrace{\dTheta=0\ , \abs{\dPhi}>0},$}
    \item{\makebox[.35\linewidth]{General \hfill} $\rBrace{\abs{\dTheta}>0\ , \abs{\dPhi}>0}.$}
\end{itemize}
\ifdefined\showTodo
{
    \subsection*{Small estimation error analysis - \textbf{TO BE REMOVED - kept only for the todos}}
    \label{subsection_ArrayPerformance_TayolrAnalysis}
    \myTodo{inline}{\textbf{DONE:}\\ maybe to call this subsection 'small error analysis'}
    Plugging (\ref{eqn_CB_coefSet}) into (\ref{eqn:GeneralFeedbackTransferFunction}) and denoting $\D{N}{x} \triangleq \frac{\sin{\rBrace{Nx}}}{\sin{\rBrace{x}}}$ results in the general \coefSetName{} beampattern \myTodo{inline}{\textbf{DONE}\\you need to define the delta expressions (the errors)}
    \begin{equation*}
        h_{\coefSetName{}}\rBrace{\theta,\omega}
        =
        \frac{
        \D{N}{\dTheta/2}exp\rBrace{-j\rBrace{\dPhi+\frac{N-1}{2}\dTheta}}
        }{
        N - \D{N}{\dTheta/2}exp\rBrace{-j\rBrace{\dPhi+\frac{N-1}{2}\dTheta}}
        }.
    \end{equation*}
    For the evaluation of the various array parameters, we define the normalized beampattern \myTodo{inline}{\textbf{DONE:}\\not very readable. Consider writing a multiplication of two terms. Suggesting that you write the expression in eq(5) as $H(\theta,\phi)$ and here you'll have $H(\hat{\theta}, \hat{\phi})/ H(\theta,\phi)$}
    \begin{equation}
        \label{eqn_arrPerformance_beamwidth_3dB}
        \Hr{\theta}{\tau}{}\triangleq\fbBpRatio.
    \end{equation}
    \myTodo{inline}{\textbf{DONE:}\\this next section is very long. Suggesting that you express (10) in terms of the errors $\Delta\theta,\;\Delta\phi$, and then simply state that the second order Taylor expansion of those errors around zero gives (12). One more issue which I think might cause us some headache is that we cannot assure that $\Delta\phi$ (you call it $\Delta\tau$) is indeed close to zero, as this term fluctuates very fast. But we will think about it later on. Maybe something with wideband signal stimulus will solve this issue.}
    The pursuit for analytic dependencies between $\Hr{\theta}{\tau}{}$, $\dTheta$ and $\dPhi$, as in \cite{van2004optimum}, lead to expressing $\Hr{\theta}{\tau}{}$'s multivariate Taylor expansion, setting $\dTheta,\dPhi$ as the variables. Simulations have shown that $2^{nd}$ Taylor expansion achieves very accurate results, thus we use it to express the array parameters. As commonly known, the Taylor expansion of a multivariate analyzable function $f\rBrace{\vx}$ around $\vx_{0}$ where $\vx \in \mathbb{R}_{M\times1},$ is 
    \begin{equation}
        \label{eqn_h_Tylor_dTheta_dTau}
        \evalat{f\rBrace{\vx}}{\vx\to\vx_{0}}=\sum_{n=0}^{\infty}\frac{1}{n!}\rBrace{\sum_{i=1}^{M}(x_{i}-x_{0i})\frac {\partial}{\partial x_i} }^n f(x_k)|_{x_k=x_{k0}},
    \end{equation}
    where $\frac{\partial}{\partial x_i}$ is the derivative operator and $i\in\left[1\hdots{}M\right]$. Reducing (\ref{eqn_h_Tylor_dTheta_dTau}) to its $2^{nd}$ form (i.e $f(x,y)=\sum_{n=0}^{\infty} \frac 1 {n!}\rBrace{x\frac {\partial}{\partial x}+y\frac {\partial}{\partial y} }^n f(x,y)|_{(x,y)=(0,0)}$), combined with the binomial formula, $(x+y)^{n}=\sum _{k=0}^{n}{\binom {n}{k}}x^{n-k}y^{k},$ multiple iterations of L'Hôpital's rule and algebraic simplification finally yields
    \begin{equation}
        \begin{split}
            \evalat{\Hr{\theta}{\tau}{}_{\coefSetName{}}}{\dTheta\to0,\dPhi\to0} \approx\ & 1 
            \\+&\frac{\binom{2}{0}}{2!}\frac{-\rBrace{N - 1}\rBrace{N-4r+2Nr+1}}{6\rBrace{r-1}^{2}}\dTheta^{2}
            \\+&\frac{\binom{2}{1}}{2!}\frac{-r\omega\rBrace{N - 1}}{\rBrace{r-1}^{2}}\dTheta\dPhi
            \\+&\frac{\binom{2}{2}}{2!}\frac{-2r\omega^{2}}{\rBrace{r-1}^{2}}\dPhi^{2}
        \end{split}
    \end{equation}
}
\else
\fi
\subsection*{The normalized beampattern}
\label{subsection_spatialIIR_normBP}
As commonly done for ULA performance analysis \cite{van2004optimum}, we analyze the normalized response (i.e., where the peak main lobe gain is set to $0_{dB}$). Henceforth, setting $\vecnot{\beta}_{\coefSetName,\text{opt}}=\vecnot{\alpha}_{\coefSetName,\text{opt}},$ we define the normalized response
\begin{equation}
    \label{eq_narmalized_pattern}
    %\resizebox{.89\linewidth}{!}{
    \begin{split}
        \HrTPr\omegaB&\triangleq
        \frac{
        \HbaFB
        }{
        H_{\vecnot{\beta}_{\coefSetName,\text{opt}},\vecnot{\alpha}_{\coefSetName,\text{opt}}}\omegaB
        }
         =
        \frac{
        \HbaFB
        }{
        r/\rBrace{1-r}
        },
        % \\
        % &=\frac{\rBrace{1-r}^{2}\Dp{\dTheta/2,N}{2}}{1+r^{2}\Dp{\dTheta/2,N}{2}-2r\D{\dTheta/2}{N}\cos{\rBrace{\dPhi+\frac{N-1}{2}\dTheta}}},
    \end{split}
    %}
\end{equation}
where the $\Delta\theta,\Delta\phi,r$ subscripts express the DOA, range and gain errors dependency, respectively.
%Also, $\abs{\HrTPr}$ is defined to be the beampattern.
Plugging \eqref{eq:SF_CB} into \eqref{eq_narmalized_pattern} yields 
\begin{equation}\label{eq_generalH}
    \resizebox{0.89\linewidth}{!}{
        \begin{split}
             \HrTPr\omegaB=
             \frac{\rBrace{1-r}\D{\dTheta/2}{N}}{\exp\rBrace{j\rBrace{\dPhi+(N-1)\dTheta/2}}-r\D{\dTheta/2}{N}}.
        \end{split}
        }
\end{equation}
Note that the known \cite{van2004optimum} normalized response of standard ULA is obtained by setting $r=0$ (i.e., cancelling the feedback) 
$$
\Hr_{\Delta\theta,\Delta\phi=0,r=0}\omegaB=
             \D{\dTheta/2}{N}\exp\rBrace{-j\rBrace{(N-1)\dTheta/2}}.
$$
Considering the steering error scenario (i.e., $\dTheta\neq0,\ \dPhi=0$) first, where 
\begin{equation}\label{eq_Hdphi0}
\Hr_{\Delta\theta,\Delta\phi=0,r}\omegaB=
             \frac{\rBrace{1-r}\D{\dTheta/2}{N}}{\exp\rBrace{j\rBrace{(N-1)\dTheta/2}}-r\D{\dTheta/2}{N}},
\end{equation}
we evaluate the FB's beamwidth, sidelobe level and directivity and compare them to those of the standard ULA.
\ifdefined\useOmega
To simplify the exposition, we shall suppress the $\omega$ dependency in the following sections (i.e. $\HrTPr \triangleq \HrTPr\omegaB$) where possible.
\fi
\vfill
\subsection*{Half power beamwidth}
\input{./Modules/arrayPerformance_beamwidth.tex}
\subsection*{Sidelobes attenuation}
\ifdefined\showDev
    \fbox{
    \begin{minipage}{0.9\linewidth}
    \textbf{development specifics}\\
    Let $f\rBrace{\dTheta} \triangleq \D{\dTheta/2}{N}\exp\rBrace{-j\rBrace{(N-1)\dTheta/2}},$ such that $\Hr_{\Delta\theta,\Delta\phi=0,r=0}\omegaB = \frac{f\rBrace{\dTheta}}{1-rf\rBrace{\dTheta}}.$ Then, we compute the beampattern's derivative and state that
    \begin{equation*}
    % \resizebox{0.9\linewidth}{!}{
        \begin{split}
            &\frac{\partial}{\partial\dTheta}\Hr_{\Delta\theta,\Delta\phi=0,r}\omegaB &\\
            &=\frac{\Dp{\dTheta/2,N}{'}\rBrace{1-r\D{\dTheta/2}{N}}-\D{\dTheta/2}{N}\rBrace{-r\Dp{\dTheta/2,N}{'}}}{\rBrace{1-r\D{\dTheta/2}{N}}^{2}}
            \\
            &=\frac{\Dp{\dTheta/2,N}{'}}{\rBrace{1-r\D{\dTheta/2}{N}}^{2}}.
        \end{split}
        % }
    \end{equation*}
    It follows that the sidelobes of the feedback-based beampattern are located at the same angles as the in the ULA case.
    \end{minipage}
    }
\else
\fi
By taking a derivative of $\Hr_{\Delta\theta,\Delta\phi=0,r}\omegaB$ with respect to $\dTheta$ it can be easily verified that the beampattern's extrema points are located exactly as in the standard ULA beampattern. Specifically, the sidelobes locations are
\begin{equation}
    \label{eqn_CB_sidelobesLocations}
    \Delta\theta_{\text{sidelobe}} = \frac{\rBrace{2m+1}\pi}{N}\ \forall m\in\cBrace{\pm 1,\pm 2,\hdots}.
\end{equation}
Our main interest is with the first sidelobe (i.e. $m=1$), therefore we evaluate \eqref{eq_Hdphi0} at $\Delta\theta = 3\pi/N$, which results in
\begin{equation}
    \abs{\Hr_{{3\pi}/{N},0,r}\omegaB}^2
    =
    \frac{
    2\rBrace{1-r}^{2}
    }{
    \rBrace{N^{2}-2Nr}\rBrace{1-\cos{\rBrace{\frac{3\pi}{N}}}}+2r^{2}
    }
    \label{eq_HSidelobes}
\end{equation}
and for large $N$ values 
\begin{equation*}
    \lim_{N\rightarrow\infty}\abs{\Hr_{3\pi/N,0,r}\omegaB}=\frac{
    2\rBrace{1-r}
    }{
    3\pi
    }.
\end{equation*}
\ifdefined\showDev
    \fbox{
    \begin{minipage}{0.9\linewidth}
    \textbf{development specifics}\\
    We know that \eqref{eq_Hdphi0} can be rewritten as
    \begin{equation*}
        \begin{split}
            \Hr_{\Delta\theta,\Delta\phi=0,r=0} \omegaB=
            \frac{
            \rBrace{1-\cos{\rBrace{N\dTheta}}}\rBrace{1-r}^{2}
            }{
            \begin{split}
                N^{2}&\rBrace{1-\cos{\dTheta}}+r^{2}\rBrace{1-\cos{N\dTheta}}
                \\
                &+Nr\Bigg(1+\cos{\rBrace{\rBrace{N-1}\dTheta}}
                \\
                &-\cos{\rBrace{N\dTheta}-\cos{\rBrace{\dTheta}}}\Bigg)
            \end{split}
            }
        \end{split}
    \end{equation*}
    Using the symbolic toolbox in MATLAB, setting $\dTheta=3\pi/N$, results in \eqref{eq_HSidelobes}.
    \end{minipage}
    }
\else
\fi
\par For standard ULA, the gain of the first sidelobe is known to be $2/3\pi$ \cite{van2004optimum}, which implies that the first sidelobe is smaller by a factor of $1-r$ compared to standard ULA.
Specifically, in perfect gain match scenario (i.e., $r\to{}1$) , the sidelobes vanish. 
\subsection*{Array directivity}
The array directivity $\mathcal{D}$ \cite{van2004optimum}, defined as
\begin{equation}\label{eq_D}
    \mathcal{D}\rBrace{N,r} = \frac{\Hr_{\Delta\theta=0,\Delta\phi=0,r}}{\frac{1}{2\pi}\int_{0}^{2\pi}\Hr_{\Delta\theta,\Delta\phi=0,r}\ d\Delta\theta} = \frac{2\pi}{\int_{0}^{2\pi}\Hr_{\Delta\theta,\Delta\phi=0,r}\ d\Delta\theta},
\end{equation}
measures the ratio between the maximal array gain at its mainlobe, to the average gain over all directions. 
For uniformly weighted ULAs with no feedback, it is known \cite{van2004optimum} that $\mathcal{D}\rBrace{N,0} = N$.
Plugging \eqref{eq_Hdphi0} within \eqref{eq_D} and by numerical evaluation (see Fig.~\ref{fig_directivity}) we suggest to approximate the directivity with 
\begin{equation}\label{eq_D_result}
    \mathcal{D}\rBrace{N,r} \approx \frac{N-r}{1-r},
\end{equation}
where the standard ULA's known result is obtained for $r=0$.
Also, for $N\geq2$, $\lim_{r\rightarrow 1}\mathcal{D}\rBrace{N,r}=\infty$, implying infinite directivity for the perfectly gain-matched FB. 
\par Finally, expressing the improvement in directivity compared to the standard ULA, assuming large $N$ values, gives rise to
\begin{equation}\label{eq_Dimprovement}
\lim_{N\to\infty}\frac{\mathcal{D}\rBrace{N,r}}{\mathcal{D}\rBrace{N,0}}
=\frac{N/\rBrace{1-r}}{N}=\frac{1}{1-r}.
\end{equation}
\begin{figure}[t]
    \begin{center}
        \begin{overpic}[width=0.75\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/directivity_expr.eps}
            \put (20, 10){\footnotesize{$N$}}
            \put (72, 6){\footnotesize{$r$}}
            \put (16.5, 70){\footnotesize{Numerical integration}}
            \put (16.5, 65){\footnotesize{$\rBrace{N-r}/\rBrace{1-r}$}}
            \put (-9, 45){\footnotesize{$\mathcal{D}\rBrace{N,r}$}}
        \end{overpic}
    \end{center}
     \caption{Plot of $\mathcal{D}\rBrace{N,r}$, computed using numerical integration (surface), shown to perfectly match the analytic expression (black diamonds) presented in \eqref{eq_D_result}. Difference between the numerical integration result and the suggested analytic expression is around $\sim{1e^{-9}}$.}
    \label{fig_directivity}
\end{figure}
\subsection*{Summary}
To conclude this section, we summarize the feedback integration related performance improvements in Table.~\ref{table_arrayPerformance} 
\begin{table}[h!]
    \caption{Performances of Classical ULA and the Proposed Feedback-Beamforming Architecture, with a Gain Mismatch $r$.}
    \centering
    \resizebox{1\linewidth}{!}{
        \begin{tabular}{||c c c c||}
            \hline
            & ULA & \thead{FEEDBACK\\BEAMFORMING} & IMPROVEMENT \\ [0.5ex] 
            \hline\hline
            HPBW & $ 1.4/N$ & $1.4/\rBrace{f(r)N}$ & $f\rBrace{r}$ times smaller\\ 
            \thead{FIRST\\SIDELOBE\\GAIN} & $2/3\pi$ & $2\rBrace{1-r}/3\pi$ & \thead{smaller by a factor of $1-r$\\for $N\gg{}1$} \\
            DIRECTIVITY & $N$ & $\rBrace{N-r}/\rBrace{1-r}$ & \thead{$1/\rBrace{1-r}$ times higher \\ for $N\gg{}1$}\\
            [1ex] 
            \hline
         \end{tabular}
     }
    \label{table_arrayPerformance}
\end{table}