### 2020-10-31 front/abbrevs.tex

% Just write down your abbervations here - or comment-out the command

\abbreviationsAndNotation{
\begin{tabular}{p{2cm}@{:\quad}l}
QED & quod erat demonstratum \\
$c$ & the speed of light \\
$a\pm b$ & the closed interval $\left[a-b,a+b\right]$ \\
\end{tabular}
}