### 2020-11-06 main/prelims.tex

\chapter{Preliminaries}
\label{chap:prelims}

In this chapter, we overview the theoretical basis of the thesis. 
In section 1...
In section 2...
In section 3...

\begin{definition}
The \emph{von Neumann model} of a computer, also known as the \emph{Princeton architecture} is an architecture for digital computers, which consists of a processing units, containing an ALU and processing registers; a control unit consisting of an instruction register and a program counter; a memory unit which stores both data and instructions; and input-and-output mechanisms.
\end{definition}

\section{Propagating Wave Fields}
\label{sec:prlm_propWaveField}
\input{Modules/sec_prlm_propWaveField}
\section{Sensor arrays}
\label{sec:prlm_sensorArrays}
\input{Modules/sec_prlm_sensorArrays}
\section{Time-Space Signals}
\label{sec:prlm_timeSpaceSig}
\input{Modules/sec_prlm_timeSpaceSig}
\section{Sensor array performance measure}
\label{sec:prlm_sensorArrayPerf}
\input{Modules/sec_prlm_arrPerf}
\section{Finite Impulse Response Based Spatial Filtering}
\label{sec:prlm_FIR}
\input{Modules/sec_prlm_FIR}
\section{Infinite Impulse Response Filters}
\label{sec:prlm_IIR}
\section{Fisher Information Matrix}
\label{sec:prlm_FIM}

Let's define some concept we'll be using throughout the thesis.