### 2021-01-08 Modules/sec_future_coef.tex

As mentioned in Sec.~\ref{sec_introduceFeedback}, the reasoning for the selection of $\vecnot{\alpha}, \vecnot{\beta}$ as the conventional beamformer generalization, was the FIM related information considerations.
Obviously, there are many more possibilities/considerations to choosing the coefficient.
Some examples for such coefficient choosing schemes may include
\begin{itemize}
    \item Finite word affects
    \item Nulling of specific DOAs
    \item Setting all poles in a single DOA may lead to enhanced directivity
    \item General radiation patterns of the sensors
\end{itemize}
and many more.