### 2021-01-08 Modules/sec_future_multiTarget.tex

Considering practical scenarios of multiple ($P$) targets, positioned in $p_{t_,i}\ \forall i\in0\dots{}P-1$, the obvious expansion of \eqref{eqn:SingleSensorTemporalEquality} is 
\begin{equation}
    \label{eqn:ftr_dyn_temp}
    x_{n}(t) = \sum_{i=0}^{i=P-1}{\rBrace{g\rBrace{s\rBrace{t-\tau_{pd,i}-\tau_{n,i}}
    +\sum_{m=0}^{N-1}{\alpha^{*}_{m}x_{m}\rBrace{t-\tau_{pd,i}-\tau_{n,i}}}}}},
\end{equation}
where $\tau_{pd,i}$ is the delay of the signal which reflected from $p_{t,i}$.
Interesting directions of research may include
\begin{itemize}
    \item The affects of closely positioned targets
    \item Localization resolution
    \item Maximal number of simultaneously detected targets with respect to N
    \item Performance of independent super-positioned dual-frequency (see Sec.~\ref{sec_app}) systems.
\end{itemize}