### 2021-02-12 front/abbrevs.tex

% Just write down your abbervations here - or comment-out the command

\abbreviationsAndNotation{
\begin{longtable}{p{2cm}@{:\quad}l}
DOA & Direction of arrival \\
RADAR & RAdio Detection And Ranging \\
FIR & Finite impulse response \\
IIR & Infinite impulse response \\
ULA & Uniform linear array \\
FIM & Fisher information matrix \\
SNR & Signal to noise ratio \\
MUSIC & Multiple signal classification (algorithm)\\
RHS & Right hand side \\
FB & Feedback beamformer \\
SF & Single frequency \\
DF & Dual frequency \\
HPBW & Half power beamwidth \\
PDF & Probability density function \\
CRLB & Cram\'er Rao lower bound \\
CRB & Cram\'er Rao bound \\
PCA & Principal component analysis \\
ICA & Independent components analysis \\
MVDR & Minimum Variance Distortionless Response \\
UWB & Ultra wide band \\
HOS & High order statistics \\
$\vecnot{v}$ & A vector \\
$v_{i}$ & The $i$-th element of vector $\vecnot{v}$ \\
$\vecnot{v}^{T}$ & Transpose of vector $\vecnot{v}$ \\
$\vecnot{v}^{*}$ & Conjugate of vector $\vecnot{v}$ \\
$\vecnot{v}^{H}$ & Hermitian (conjugated) transpose of vector $\vecnot{v}$ \\
$\vecnot{A}$ & A Matrix \\
$A_{i,j}$ & The $i,j$-th element of matrix $\vecnot{A}$ \\
$\vecnot{A}^{T}$ & Transpose of matrix $\vecnot{A}$ \\
$\vecnot{A}^{*}$ & Conjugate of matrix $\vecnot{A}$ \\
$\vecnot{A}^{H}$ & Hermitian (conjugated) transpose of matrix $\vecnot{A}$ \\
$N$ & Number of elements in array \\
$d$ & Array's inter-element spacing \\
$\omega$ & Radial frequency \\
$c$ & Wave's propagation velocity \\
$\delta{}x$ & Small diff in $x$ \\
$\lambda$ & Wavelength \\
$t$ & Time \\
$\theta_{d}$ & Geometrical angle \\
$\psi$, $\theta$ & Electrical phase \\
\vecnot{d} & Steering vector \\
$B$ & beampattern \\
$R$ & Distance to target \\
$z\rBrace{t}$ & Array's temporal output \\
$Z\rBrace{\omega}$ & Array's frequency-domain output \\
$\mathcal{D}$ & Directivity \\
$\mathcal{H}$ & Normalized beampattern \\
\end{longtable}
}