### 2021-02-12 main/chap_futureResearch.tex

In the following chapter, some of the many possibilities for further investigation of the feedback beamformer are suggested.
As mentioned in Sec.~\ref{sec_introduceFeedback}, we assumed a single target stationary scenario where the target's location is fixed.
Naturally, the first suggested subjects for future research are dynamic (Sec.\ref{sec_future_dynamic}) and multiple (Sec.\ref{sec_future_multiTarget}) target scenarios which obviously should be followed by a combined general scenario of multiple dynamic targets performance scenario.
Another setting which may be an interesting subject for future research is the waveform characteristics (Sec.~\ref{sec_future_wavfrom}) and it's affect on the overall spatial performance.
We choose to conclude with another final suggestion (Sec.~\ref{sec_future_coef}) - i.e. to revisit Sec.\ref{sec_FIM}'s FIM related considerations to the choice of array settings. 
\subsection{Dynamic target}
\label{sec_future_dynamic}
\input{Modules/sec_future_dynamic}
\subsection{Multi-target scenario}
\label{sec_future_multiTarget}
\input{Modules/sec_future_multiTarget}
\subsection{Waveform modifications}
\label{sec_future_wavfrom}
\input{Modules/sec_future_wavfrom}
\subsection{Coefficients quality criteria}
\label{sec_future_coef}
\input{Modules/sec_future_coef}
% \section{Apply to other applications}
% \label{sec_future_applications}
% \input{Modules/sec_future_applications}