### 2021-02-12 main/prelims.tex

\chapter{Preliminaries}
\label{chap:prelims}
As a preliminary discussion to the presentation of our contribution, we overview some the key theoretical basis of the thesis.
First, a short overview of array-processing is presented in Sec.~\ref{sec:prlm_sensorArrays}, accompanied by a complementary wave propagation discussion in Sec.~\ref{sec:prlm_propWaveField}.
To accommodate the discussion for spatial filtering, we also mention the time-space relation of spatially sampled signals in Sec.~\ref{sec:prlm_timeSpaceSig}.
In Sec.~\ref{sec:prlm_localization}, we further focus the spatial filtering discussion to localization and overview some of its fundamentals.
Following the FIR/IIR filters related discussion in Chapter.~\ref{chap:intro}, we further elaborate and compare the two architectures in Sec.~\ref{sec:prlm_FIR_IIR}.
To complete the basics overview, we also discuss array performance measuring in Sec.~\ref{sec:prlm_array_perf} following the classical localization related literature.
% \begin{definition}
% The \emph{von Neumann model} of a computer, also known as the \emph{Princeton architecture} is an architecture for digital computers, which consists of a processing units, containing an ALU and processing registers; a control unit consisting of an instruction register and a program counter; a memory unit which stores both data and instructions; and input-and-output mechanisms.
% \end{definition}

% \section{Digital Filter Design - Finite/Infinite Response Filters}
% \label{sec:prlm_FILTERS}
% \input{Modules/sec_prlm_FILTERS}
\section{Sensor arrays}
\label{sec:prlm_sensorArrays}
\input{Modules/sec_prlm_sensorArrays}
\section{Propagating Wave Fields}
\label{sec:prlm_propWaveField}
\input{Modules/sec_prlm_propWaveField}
\section{Spatio-temporal processing}
\label{sec:prlm_timeSpaceSig}
\input{Modules/sec_prlm_timeSpaceSig}
\section{Localization}
\label{sec:prlm_localization}
\input{Modules/sec_prlm_localization}
% \section{Sensor array performance measure}
% \label{sec:prlm_sensorArrayPerf}
% \input{Modules/sec_prlm_arrPerf}
\section{The analogy between array processing and temporal digital filtering}
\label{sec:prlm_FIR_IIR}
\input{Modules/sec_prlm_FIR_IIR}
\section{Array spatial performance}
\label{sec:prlm_array_perf}
\input{Modules/sec_prlm_array_perf}