### 2021-05-14 Modules/paper_sec_prlms.tex

In this chapter, we revisit the most basic DOA estimation approach, i.e. beamforming, and present our novel feedback based architecture.
The feedback based beamforming concept is elaborated in Sec.~\ref{sec_introduceFeedback}.
Then, we apply information theory related reasoning for the array setting in Sec.~\ref{sec_FIM}, namely we use the FIM to design the coefficients of the suggested beamformer.
The resultant beamformer's temporal response and spatial performance are analysed in Sec.~\ref{sec_stability} and Sec.~\ref{sec_Performance} respectively, assuming ideal noiseless scenario.
Considering noisy setups, the beamformer's noise sensitivity is exemplified in Sec.~\ref{sec_sim}, followed by a practical and robust dual-frequency beamformer design which is presented in Sec.~\ref{sec_app}.

\section{Notations and problem setup}
\label{sec_introduceFeedback}
In this section, a feedback-based architecture is proposed for spatial signal processing.
Inspired by time domain ``Direct form II'' IIR filter design (see Fig.~\ref{fig_IIR_arch}), 
we propose to use the same concept in the spatial domain.
The suggested FB architecture, where the output signal ($z$) is synthesized using weights $\vBeta$ and the weights $\vAlpha$ synthesize the feedback transmission ($\Tx$), is presented in Fig.~\ref{fig:Proposed_spatialIIR_ARCH}. The beamformer's output and the feedback signal are synthesized using two independently configured beamformers, $s$ is the system's stimulus and an additive noise (n) is assumed at the array's output.
Also, the FB block is marked (dashed line) for later use.
Note that setting $\vAlpha=\vecnot{0}$ (i.e., cancelling the feedback) degenerates the system to the convectional beamformer (CB) beamformer.
\begin{figure}[t!]
    \begin{center}
        \begin{overpic}[width=0.5\linewidth, 
        % grid, 
        tics=10,trim={0 0 0 0}]{./Media/SpatialIIR-diagram/SpatialIIR_VER9.png}
            \put (14, 39){\footnotesize{$\sum_{n=0}^{N-1}\beta^{*}_{n}x_{n}$}}
            \put (54, 39){\footnotesize{$\sum_{n=0}^{N-1}\alpha^{*}_{n}x_{n}$}}
            \put (24.25, 50){\footnotesize{$x_{0}$}}
            \put (35, 50){\footnotesize{$x_{1}$}}
            \put (55, 50){\footnotesize{$x_{N-1}$}}
            \put (30.5, 66){\footnotesize{$\delta$}}
            \put (89, 96){\footnotesize{$\vpt$}}
            \put (58, 58){\footnotesize{$\vpi{N-1}$}}
            \put (37, 58){\footnotesize{$\vpi{1}$}}
            \put (26.5, 58){\footnotesize{$\vpi{0}$}}
            \put (41.5, 64.5){\footnotesize{$\theta_{d}$}}
            \put (19, 11.25){\large{$+$}}
            \put (61.25, 27.25){\large{$+$}}
            \put (47, 11.75){$s$}
            \put (35.25, 11.75){$n$}
            \put (21,4){$z$}
            \put (63.5,14){$\Tx$}
            \put (1, 51){$\text{FB}_{\vAlpha,\vBeta}$}
        \end{overpic}
    \end{center}
    \caption{
    The proposed FB.
    The spatial feedback is obtained by continuous re-transmission of $\Tx$ to the target at $\vpt$.
    We designate the FB block (dashed line) for later use.
    }
    \label{fig:Proposed_spatialIIR_ARCH}
\end{figure}
\subsection{Obtained spatial response}
\input{./Modules/Spatial_IIR-MathDerivations.tex}
\section{Fisher Information Matrix}
\label{sec_FIM}
\input{./Modules/sec_FIM_and_CRLB.tex}
\section{Temporal Stability}
\label{sec_stability}
\input{./Modules/sec_stability.tex}
\section{Performance Analysis}
\label{sec_Performance}
In this section we analyze the suggested FB (see Fig.~\ref{fig:Proposed_spatialIIR_ARCH}), considering some fundamental properties which are commonly used to asses array performance and were discussed in Sec.~\ref{sec:prlm_array_perf}: Beamwidth, peak-to-sidelobe level, and directivity. Each property is then compared to traditional passive ULAs, showing that significantly improved performances are obtainable with spatial feedback integration.
\subsection{The normalized beampattern}
\label{subsection_spatialIIR_normBP}
Applying the \emph{normalized beampattern} notation of Sec.~\ref{sec:prlm_array_perf}, we set $\vecnot{\beta}_{\coefSetName}=\vecnot{\alpha}_{\coefSetName}.$ 
Considering the unavoidable errors mentioned in \ref{subsec_error_trms} (i.e. $\Delta\theta, \Delta\phi$ and $r$), we define $\HrTPr$ as the general scenario normalized array response.
We choose to normalize the spatial response according to $H_{0}$ (of \eqref{eq:SF_CB_gainMismatch}), giving rise to
\begin{equation}
    \label{eq_narmalized_pattern}
    %\resizebox{.89\linewidth}{!}{
    \begin{split}
        \HrTPr&\triangleq
        \frac{
        \HbaFB
        }{
        H_{0}
        }\\
        &=
        \frac{
        \HbaFB
        }{
        r/\rBrace{1-r}
        }\\
        &=
        \frac{\rBrace{1-r}\D{\dTheta/2}{N}}{\exp\cBrace{j\vBrace{\dPhi+\rBrace{N-1}\dTheta/2}}-r\D{\dTheta/2}{N}}.
    \end{split}
\end{equation}
Note that the known \cite{van2004optimum} normalized response of standard ULA is obtained by setting $r=0$ and $\Delta\phi=0$
$$
\Hr_{\Delta\theta,\Delta\phi=0,r=0}=
             \D{\dTheta/2}{N}\exp\rBrace{-j\rBrace{(N-1)\dTheta/2}}.
$$
Considering the steering error scenario (i.e., $\dTheta\neq0,\ \dPhi=0$) first, where 
\begin{equation}\label{eq_Hdphi0}
\Hr_{\Delta\theta,\Delta\phi=0,r}=
             \frac{\rBrace{1-r}\D{\dTheta/2}{N}}{\exp\cBrace{j\vBrace{\rBrace{N-1}\dTheta/2}}-r\D{\dTheta/2}{N}},
\end{equation}
we evaluate the FB's beamwidth, sidelobe level and directivity and compare them to those of the standard ULA.
\subsection{Half power beamwidth}
\input{./Modules/arrayPerformance_beamwidth_LOCAL.tex}
\subsection{Sidelobes attenuation}
\ifdefined\showDev
    \fbox{
    \begin{minipage}{0.9\linewidth}
    \textbf{development specifics}\\
    Let $f\rBrace{\dTheta} \triangleq \D{\dTheta/2}{N}\exp\rBrace{-j\rBrace{(N-1)\dTheta/2}},$ such that $\Hr_{\Delta\theta,\Delta\phi=0,r=0} = \frac{f\rBrace{\dTheta}}{1-rf\rBrace{\dTheta}}.$ Then, we compute the beampattern's derivative and state that
    \begin{equation*}
    % \resizebox{0.9\linewidth}{!}{
        \begin{split}
            &\frac{\partial}{\partial\dTheta}\Hr_{\Delta\theta,\Delta\phi=0,r} &\\
            &=\frac{\Dp{\dTheta/2,N}{'}\rBrace{1-r\D{\dTheta/2}{N}}-\D{\dTheta/2}{N}\rBrace{-r\Dp{\dTheta/2,N}{'}}}{\rBrace{1-r\D{\dTheta/2}{N}}^{2}}
            \\
            &=\frac{\Dp{\dTheta/2,N}{'}}{\rBrace{1-r\D{\dTheta/2}{N}}^{2}}.
        \end{split}
        % }
    \end{equation*}
    It follows that the sidelobes of the feedback-based beampattern are located at the same angles as the in the ULA case.
    \end{minipage}
    }
\else
\fi
By taking a derivative of $\Hr_{\Delta\theta,\Delta\phi=0,r}$ with respect to $\dTheta$ it can be easily verified that the beampattern's extrema points are located exactly as in the standard ULA beampattern. Specifically, the sidelobes locations are
\begin{equation}
    \label{eqn_CB_sidelobesLocations}
    \Delta\theta_{\text{sidelobe}} = \frac{\rBrace{2m+1}\pi}{N}\ \forall m\in\cBrace{\pm 1,\pm 2,\hdots}.
\end{equation}
Our main interest is with the first sidelobe (i.e. $m=1$), therefore we evaluate \eqref{eq_Hdphi0} at $\Delta\theta = 3\pi/N$, which results in
\begin{equation}
    \abs{\Hr_{{3\pi}/{N},0,r}}^2
    =
    \frac{
    2\rBrace{1-r}^{2}
    }{
    \rBrace{N^{2}-2Nr}\rBrace{1-\cos{\rBrace{\frac{3\pi}{N}}}}+2r^{2}
    }
    \label{eq_HSidelobes}
\end{equation}
and for large $N$ values 
\begin{equation*}
    \lim_{N\rightarrow\infty}\abs{\Hr_{3\pi/N,0,r}}=\frac{
    2\rBrace{1-r}
    }{
    3\pi
    }.
\end{equation*}
\ifdefined\showDev
    \fbox{
    \begin{minipage}{0.9\linewidth}
    \textbf{development specifics}\\
    We know that \eqref{eq_Hdphi0} can be rewritten as
    \begin{equation*}
        \begin{split}
            \Hr_{\Delta\theta,\Delta\phi=0,r=0} =
            \frac{
            \rBrace{1-\cos{\rBrace{N\dTheta}}}\rBrace{1-r}^{2}
            }{
            \begin{split}
                N^{2}&\rBrace{1-\cos{\dTheta}}+r^{2}\rBrace{1-\cos{N\dTheta}}
                \\
                &+Nr\Bigg(1+\cos{\rBrace{\rBrace{N-1}\dTheta}}
                \\
                &-\cos{\rBrace{N\dTheta}-\cos{\rBrace{\dTheta}}}\Bigg)
            \end{split}
            }
        \end{split}
    \end{equation*}
    Using the symbolic toolbox in MATLAB, setting $\dTheta=3\pi/N$, results in \eqref{eq_HSidelobes}.
    \end{minipage}
    }
\else
\fi
\par For standard ULA, the gain of the first sidelobe is known to be $2/3\pi$ \cite{van2004optimum}, which implies that the first sidelobe is smaller by a factor of $1-r$ compared to standard ULA.
Specifically, in perfect gain match scenario (i.e., $r\to{}1$) , the sidelobes vanish. 
\subsection{Array directivity}
In the context of this work, the directivity is expressed as
\begin{equation}\label{eq_D}
    \mathcal{D}\rBrace{N,r} = \frac{\Hr_{\Delta\theta=0,\Delta\phi=0,r}}{\frac{1}{2\pi}\int_{0}^{2\pi}\Hr_{\Delta\theta,\Delta\phi=0,r}\ d\Delta\theta} = \frac{2\pi}{\int_{0}^{2\pi}\Hr_{\Delta\theta,\Delta\phi=0,r}\ d\Delta\theta},
\end{equation}
Plugging \eqref{eq_Hdphi0} within \eqref{eq_D} and by numerical evaluation (see App.~\ref{apndx_directivityFit} and the complementary Fig.~\ref{fig_directivity}) we suggest to approximate the directivity with 
\begin{equation}\label{eq_D_result}
    \mathcal{D}\rBrace{N,r} \approx \frac{N-r}{1-r},
\end{equation}
where the standard ULA's known result is obtained for $r=0$.
Also, for $N\geq2$, $\lim_{r\rightarrow 1}\mathcal{D}\rBrace{N,r}=\infty$, implying infinite directivity for the perfectly gain-matched FB. 
\par Finally, expressing the improvement in directivity compared to the standard ULA, assuming large $N$ values, gives rise to
\begin{equation}\label{eq_Dimprovement}
\lim_{N\to\infty}\frac{\mathcal{D}\rBrace{N,r}}{\mathcal{D}\rBrace{N,0}}
=\frac{N/\rBrace{1-r}}{N}=\frac{1}{1-r}.
\end{equation}
\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \begin{center}
            \begin{overpic}[width=\linewidth, 
                %grid, 
                tics=10,trim=0 0 0 0]{./Media/directivity_expr.eps}
                    \put (20, 10){\footnotesize{$N$}}
                    \put (72, 6){\footnotesize{$r$}}
                    \put (16.5, 70){\footnotesize{Numerical integration}}
                    \put (16.5, 65){\footnotesize{$\rBrace{N-r}/\rBrace{1-r}$}}
                    \put (-9, 45){\footnotesize{$\mathcal{D}\rBrace{N,r}$}}
            \end{overpic}
        \end{center}
         \caption{Plot of $\mathcal{D}\rBrace{N,r}$, computed using numerical integration (surface), shown to perfectly match the analytic expression (black diamonds) presented in \eqref{eq_D_result}.}
        \label{fig_directivity}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \begin{center}
            \begin{overpic}[width=\linewidth, 
                % grid, 
                tics=10,trim=0 0 0 0]{./Media/directivityErr.png}
                    \put (18, 6){\footnotesize{$N$}}
                    \put (72, 3){\footnotesize{$r$}}
                    \put (0, 68){\footnotesize{Error}}
            \end{overpic}
        \end{center}
        \caption{The error between numerical calculation and the analytic expression.}
        \label{fig_directivityErr}
    \end{minipage}
\end{figure}
% \begin{figure}[t]
%     \begin{center}
%         \begin{overpic}[width=0.75\linewidth, 
%         %grid, 
%         tics=10,trim=0 0 0 0]{./Media/directivity_expr.eps}
%             \put (20, 10){\footnotesize{$N$}}
%             \put (72, 6){\footnotesize{$r$}}
%             \put (16.5, 70){\footnotesize{Numerical integration}}
%             \put (16.5, 65){\footnotesize{$\rBrace{N-r}/\rBrace{1-r}$}}
%             \put (-9, 45){\footnotesize{$\mathcal{D}\rBrace{N,r}$}}
%         \end{overpic}
%     \end{center}
%      \caption{Plot of $\mathcal{D}\rBrace{N,r}$, computed using numerical integration (surface), shown to perfectly match the analytic expression (black diamonds) presented in \eqref{eq_D_result}.}
%     \label{fig_directivity}
% \end{figure}
\subsection{Summary}
To conclude this section, we summarize the feedback integration related performance improvements in Table.~\ref{table_arrayPerformance}.
\begin{table}[h!]
    \caption{Performances of Classical ULA and the Proposed Feedback-Beamforming Architecture, with a Gain Mismatch $r$.}
    \centering
    %\resizebox{1\linewidth}{!}
    {
        \begin{tabular}{||c c c c||}
            \hline
            & ULA & \thead{FEEDBACK\\BEAMFORMING} & IMPROVEMENT \\ [0.5ex] 
            \hline\hline
            HPBW & $ 1.4/N$ & $1.4/\rBrace{f(r)N}$ & Narrower by a factor of $f\rBrace{r}$\\ 
            \thead{FIRST\\SIDELOBE\\GAIN} & $2/3\pi$ & $2\rBrace{1-r}/3\pi$ & \thead{smaller by a factor of $1-r$\\for $N\gg{}1$} \\
            DIRECTIVITY & $N$ & $\rBrace{N-r}/\rBrace{1-r}$ & \thead{$1/\rBrace{1-r}$ times higher \\ for $N\gg{}1$}\\
            [1ex] 
            \hline
         \end{tabular}
     }
    \label{table_arrayPerformance}
\end{table}
Also, in Fig.~\ref{fig_perf} we demonstrate the HPBW and sidelobe attenuation for 3 elements ULA.
As the expressions for HPBW and sidelobe-attenuation are relevant for $\sim{}N>20$, we also simulate 20 elements ULA in Fig.~\ref{fig_perf20}. 
In Table.~\ref{table_arrayPerfEmp} we show the coherency between the results and the performance related expressions of Table.~\ref{table_arrayPerformance}.
As predicted, the results of Fig.~\ref{fig_perf20} are consistent with the theoretical expressions.
\begin{figure}[t]
    \begin{center}
        \begin{overpic}[width=0.9\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/fig_performance.png}
            \put (9, 36.5){$dB$}
            \put (2, 20){$\abs{H}^{2}$}
            \put (27, 0){$\theta_{d}/\pi$}
            \put (34.5, 67){$r=0$}
            \put (29, 63){\rotatebox{-90}{\tiny{HPBW = $0.1\pi$}}}
            \put (40, 61){\tiny{$-19.08_{dB}$}}
            \put (77.5, 67){$r=0.3$}
            \put (73.14, 63){\rotatebox{-90}{\tiny{HPBW = $0.073\pi$}}}
            \put (83.5, 58.5){\tiny{$-23.43_{dB}$}}
            \put (34.5, 31.5){$r=0.6$}
            \put (29, 27){\rotatebox{-90}{\tiny{HPBW = $0.043\pi$}}}
            \put (40, 18.75){\tiny{$-31.09_{dB}$}}
            \put (77.5, 31.5){$r=0.8$}
            \put (73.14, 27){\rotatebox{-90}{\tiny{HPBW = $0.02\pi$}}}
            \put (84.25, 13.5){\tiny{$-41.6_{dB}$}}
        \end{overpic}
    \end{center}
     \caption{Simulating 3 elements ULA based FB for $r$ values of 0, 0.3, 0.6 and 0.8.
     The HPBW is marked with vertical red dashed line, where an auxiliary horizontal line of $\abs{H}^{2} = 1/2$ is also provided.
     The sidelobe attenuation is cited in each plot in a textual fashion.}
    \label{fig_perf}
\end{figure}
\begin{figure}[t]
    \begin{center}
        \begin{overpic}[width=0.9\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/fig_performance20.png}
            \put (9, 36.5){$dB$}
            \put (-3, 20){$\abs{H}^{2}$}
            \put (27, 0){$\theta_{d}/\pi$}
            \put (34.5, 67){$r=0$}
            \put (29, 63){\rotatebox{-90}{\tiny{HPBW = $0.0605\pi$}}}
            \put (32.25, 56.5){\tiny{$-26.38_{dB}$}}
            \put (77.5, 67){$r=0.3$}
            \put (73, 63){\rotatebox{-90}{\tiny{HPBW = $0.0454\pi$}}}
            \put (79, 53.75){\tiny{$-32.06_{dB}$}}
            \put (34.5, 31.25){$r=0.6$}
            \put (29, 27){\rotatebox{-90}{\tiny{HPBW = $0.0259\pi$}}}
            \put (40, 14){\tiny{$-41.21_{dB}$}}
            \put (77.5, 31.25){$r=0.8$}
            \put (73, 27){\rotatebox{-90}{\tiny{HPBW = $0.0127\pi$}}}
            \put (84.25, 12.5){\tiny{$-52.55_{dB}$}}
        \end{overpic}
    \end{center}
     \caption{Simulating 20 elements ULA for the sake of HPBW analysis.
     As in Fig.~\ref{fig_perf}, $r$ values of 0, 0.3, 0.6 and 0.8 are simulated.
     The HPBW is marked in the same manner also.}
    \label{fig_perf20}
\end{figure}
\begin{table}[h!]
    \caption{Performances of CB $r=0$ and the proposed Feedback-Beamforming Architecture, for several $r$ values.
    Also comparison to expected theoretical expressions is provided.}
    \centering
    %\resizebox{1\linewidth}{!}
    {
        \begin{tabular}{||c | c c c | c c c||}
            \hline
            & \multicolumn{3}{c|}{HPBW [RAD]} & \multicolumn{3}{c||}{SIDELOBE GAIN [dB]} \\ [0.5ex]
            \hline
            Expected & \multicolumn{3}{c|}{$\frac{\rBrace{1-r}\rBrace{-0.4r+1.4}}{1.4}$} & \multicolumn{3}{c||}{$20\log{}\rBrace{\abs{\frac{2\rBrace{1-r}}{3\pi}}^{2}}$} \\ [0.5ex]
            \hline
            & Result & Expected & Error & Result & Expected & Error \\ [0.5ex] 
            \hline\hline
            $r=0$ & 0.0605 & 0.07 & 0.0095 & -26.38 & -26.93 & 0.55 \\ [0.5ex]
            $r=0.3$ & 0.0454 & 0.0448 & 0.006 & -32.06 & -33.12 & 1.06 \\ [0.5ex]
            $r=0.6$ & 0.0259 & 0.0232 & 0.0027 & -41.21 & -42.84 & 1.63 \\ [0.5ex]
            $r=0.8$ & 0.0127 & 0.0108 & 0.0019 & -52.55 & -54.88 & 2.33 \\ [0.5ex]
            \hline
         \end{tabular}
     }
    \label{table_arrayPerfEmp}
\end{table}