### 2021-05-14 Modules/sec_future_coef.tex

As mentioned in Sec.~\ref{sec_introduceFeedback}, the reasoning for the selection of $\vecnot{\alpha}, \vecnot{\beta}$ as the CB generalization, was the FIM related information considerations.
Obviously, there are many more possibilities/considerations to choosing the coefficient.
Some examples for such coefficient choosing schemes may include
\begin{itemize}
    \item Finite word effects
    \item Nulling of specific DOAs \\
    For example, it is interesting to see the FB equivalent to Capon's \cite{capon1969high} MVDR. 
\end{itemize}
and many more.