### 2021-05-14 front/abbrevs.tex

% Just write down your abbervations here - or comment-out the command

\newacronym{DOA}{DOA}{direction of arrival}
\newacronym{RADAR}{RADAR}{dadio Detection And Ranging}
\newacronym{FIR}{FIR}{finite impulse response}
\newacronym{IIR}{IIR}{infinite impulse response}
\newacronym{ULA}{ULA}{uniform linear array}
\newacronym{FIM}{FIM}{fisher information matrix}
\newacronym{SNR}{SNR}{signal to noise ratio}
\newacronym{MUSIC}{MUSIC}{multiple signal classification}
\newacronym{RHS}{RHS}{right hand side}
\newacronym{FB}{FB}{feedback beamformer}
\newacronym{SF}{SF}{single frequency}
\newacronym{DF}{DF}{dual frequency}
\newacronym{HPBW}{HPBW}{half power beamwidth}
\newacronym{PDF}{PDF}{probability density function}
\newacronym{CRLB}{CRLB}{cram\'er Rao lower bound}
\newacronym{CRB}{CRB}{cram\'er Rao bound}
\newacronym{PCA}{PCA}{principal component analysis}
\newacronym{ICA}{ICA}{independent components analysis}
\newacronym{MVDR}{MVDR}{minimum Variance Distortionless Response}
\newacronym{UWB}{UWB}{ultra wide band}
\newacronym{HOS}{HOS}{high order statistics}

\newglossaryentry{VVEC}{name=$\vecnot{v}$,description={A vector}}
\newglossaryentry{VI}{name=$v_{i}$,description={The $i$-th element of vector $\vecnot{v}$}}
\newglossaryentry{VT}{name=$\vecnot{v}^{T}$,description={Transpose of vector $\vecnot{v}$}}
\newglossaryentry{VCONJ}{name=$\vecnot{v}^{*}$,description={Conjugate of vector $\vecnot{v}$}}
\newglossaryentry{VH}{name=$\vecnot{v}^{H}$,description={Hermitian (conjugated) transpose of vector $\vecnot{v}$}}
\newglossaryentry{A}{name=$\vecnot{A}$,description={A Matrix}}
\newglossaryentry{AIJ}{name=$A_{i,j}$,description={The $i,j$-th element of matrix $\vecnot{A}$}}
\newglossaryentry{AT}{name=$\vecnot{A}^{T}$,description={Transpose of matrix $\vecnot{A}$}}
\newglossaryentry{ACONJ}{name=$\vecnot{A}^{*}$,description={Conjugate of matrix $\vecnot{A}$}}
\newglossaryentry{AH}{name=$\vecnot{A}^{H}$,description={Hermitian (conjugated) transpose of matrix $\vecnot{A}$}}
\newglossaryentry{DINTERARRAY}{name=$d$,description={Array's inter-element spacing}}
\newglossaryentry{OMEGA}{name=$\omega$,description={Radial frequency}}
\newglossaryentry{C}{name=$c$,description={Wave's propagation velocity}}
\newglossaryentry{LAMBDA}{name=$\lambda$,description={Wavelength}}
\newglossaryentry{THETAD}{name=$\theta_{d}$,description={Geometrical angle}}
\newglossaryentry{THETA}{name=$\theta$,description={Electrical phase}}
\newglossaryentry{DSTEER}{name=\vecnot{d},description={Steering vector}}
\newglossaryentry{BP}{name=$B$,description={beampattern}}
\newglossaryentry{RANGE}{name=$R$,description={Distance to target}}
\newglossaryentry{ZT}{name=$z\rBrace{t}$,description={Array's temporal output}}
\newglossaryentry{ZW}{name=$Z\rBrace{\omega}$,description={Array's frequency-domain output}}
\newglossaryentry{D}{name=$\mathcal{D}$,description={Directivity}}
\newglossaryentry{HNORM}{name=$\mathcal{H}$,description={Normalized beampattern}}