### 2021-05-14 main/prelims.tex

\chapter{Preliminaries}
\label{chap:prelims}
In this chapter, we discuss the theoretical preliminaries for this thesis.
First, a short overview of array-processing is presented in Sec.~\ref{sec:prlm_sensorArrays}, accompanied by a complementary wave propagation elementaries in Sec.~\ref{sec:prlm_propWaveField}.
As a preface to spatial filtering, we mention the time-space relation of spatially sampled signals in Sec.~\ref{sec:prlm_timeSpaceSig}, which is followed by a discussion of classic spatial performance analysis methods in Sec.~\ref{sec:prlm_array_perf} and further elaboration on
localization in Sec.~\ref{sec:prlm_localization}.
Finally, Sec.~\ref{sec:prlm_FIR_IIR} reviews the FIR/IIR filters, presents the pros and cons of both architectures and serves as the foundation for this research's motivation.
\section{Sensor arrays}
\label{sec:prlm_sensorArrays}
\input{Modules/sec_prlm_sensorArrays}
\section{Propagating Wave Fields}
\label{sec:prlm_propWaveField}
\input{Modules/sec_prlm_propWaveField}
\section{Spatio-temporal processing}
\label{sec:prlm_timeSpaceSig}
\input{Modules/sec_prlm_timeSpaceSig}
\section{Array spatial performance}
\label{sec:prlm_array_perf}
\input{Modules/sec_prlm_array_perf}
\section{Localization}
\label{sec:prlm_localization}
\input{Modules/sec_prlm_localization}
\section{The analogy between array processing and temporal digital filtering}
\label{sec:prlm_FIR_IIR}
\input{Modules/sec_prlm_FIR_IIR}